Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints\thanks{This research was supported by the Russian Science Foundation, project No. 16-11-10150.}

### 2020-02-02 2019/NUMTA_Barkalov/SoftComputing/article.tex
Renamed illustrations: Name your figure files with Fig and the figure number
correct some references
Changed some formulas

\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{lineno,hyperref,amsmath,array}
\usepackage{multirow}
\usepackage[round]{natbib} 
\usepackage[misc]{ifsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

@@

\begin{abstract}
This paper considers the constrained global optimization problems, in which the functions are of the “black box” type and satisfy the Lipschitz condition. The algorithms for solving the problems of this class require the use of adequate estimates of the a priori unknown Lipschitz constants for the problem functions. A novel approach presented in this paper is based on a simultaneous use of two estimates of the Lipschitz constant: an overestimated and an underestimated one. The upper estimate provides the global convergence whereas the lower one reduces the number of trials necessary to find the global optimizer with the required accuracy. The considered algorithm for solving the constrained problems doesn’t use the ideas of the penalty function method; each constraint of the problem is accounted for separately. The convergence conditions of the proposed algorithm are formulated in the corresponding theorem. The results of the numerical experiments on a series of multiextremal problems with non-convex constraints demonstrating the efficiency of the proposed scheme of dual Lipschitz constant estimates are presented.
\keywords{Global optimization \and Multiextremal problems \and Non-convex constraints \and Lipschitz constant estimates \and Dimensionality reduction \and Numerical methods}
\subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

	It is supposed that the objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_{j}(y), 1 \leq j \leq m$, of the constraints satisfy the Lipschitz condition
\begin{align}\label{lipschitz_condition}
	\left|g_j(y')-g_j (y'')\right| \leq L_j \|y'-y''\|, \nonumber \\ 
	y', y''\in D, \; 1\leq j \leq m+1,
\end{align}
with a priori unknown constants $L_j, 1\leq j \leq m+1$, and may be multiextremal.

@@

\begin{figure}
	\center{\includegraphics[width=0.6\linewidth]{epsilon_reserved_solution}}
	\caption{$\varepsilon$-reserved solution.}
	\label{fig:eps_reserved_solution}
\end{figure}

@@

	We will consider the initial problem (\ref{problem}) assuming that the problem functions $g_{j}(y), 1 \leq j \leq m+1,$ may be defined only partly. It means that they are defined and computable only in the sets 
\begin{align}\label{D_sets}
	D_1 = D, \; D_{j+1} = \left\{ y \in D_j: \; g_j(y) \leq 0 \right\}, \nonumber \\
	1 \leq j \leq m.
\end{align}
This situation is typical for the optimal design problems when some conditions for the functioning of the modeled object are not met and its other characteristics appear to be undefined. In this approach, the first violation of any constraint indicating that the point $y$ being considered is no longer feasible terminates further computations at the point $y$.

	Many well known approaches to solving the multidimensional global optimization problems are based on either explicit or implicit reduction of multidimensional problems to one-dimensional ones (see, for example, the description of methods using the diagonal partitions \cite{Sergeyev2006} or simplicial partitions \cite{Zilinskas2008}). In this paper, we will use the approach based on the dimensionality reduction with the use of the space-filling curve $y(x)$ (the Peano curve). Continuous unambiguous mapping $y(x)$ projecting the interval [0,1] onto an $N$-dimensional hypercube $D$ allows reducing a multidimensional constrained minimization problem in the domain $D$ to a one-dimensional constrained optimization problem in the interval [0,1]
\begin{align}\label{one_dimensional_problem}
	& \varphi(y(x^\ast)) =  \nonumber \\
	& \min \left\{\varphi(y(x)): x \in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\}.
\end{align}
The issues of the numerical construction of the Peano space filling curve-type mappings and the corresponding theory were described in detail in \cite{Strongin2000,Sergeyev2013}. Note that the theoretical Peano curve $y(x)$ is a limit object and in practice is replaced by its approximation (called \textit{evolvent}) with predefined precision $2^{-m}$, where the number $m>1$ is a parameter (the \textit{density} of the evolvent). The evolvent density is determined by the accuracy of the possible practical implementation of the problem solution found. For  illustration, Fig. \ref{fig:evolvent} (a), (b) shows the projections of the interval [0,1] for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
\begin{figure}
	\begin{minipage}[h]{0.45\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=4} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.45\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{evolvent_N=2_m=5} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Evolvent for $N=2$ and $m=4$ (a), $m=5$ (b)}
	\label{fig:evolvent}
\end{figure}

@@

	The convergence conditions for the index global search algorithm can be formulated in the form of the following theorem from \cite{Strongin2000}. 
\\
\begin{theorem}\label{theorem:1} Assume that the following conditions are satisfied:
	\begin{enumerate} 
		\item Problem (\ref{problem}) has an $\varepsilon$-reserved solution $y_{\varepsilon}$ from (\ref{epsilon_reserved_solution}).
		\item Functions $g_j(y), 1 \leq j \leq m+1,$ admit Lipschitzian (with the constants $L_j$) extensions $G_{j}(y)$ over the whole feasible domain $D$ from (\ref{D}), i.e.
	$$
		g_j \left( y(x) \right) = G_j \left( y(x) \right), \; x \in Q_j, \; 1 \leq j \leq m+1,
	$$
	where the sets $Q_j$ are from (\ref{Q_intervals}), and $y(x)$ is the curve from (\ref{one_dimensional_problem}).
		\item The parameters of the method $\epsilon_{\nu}, 1 \leq \nu \leq m,$ used in the rule (\ref{z_estimates}) are the corresponding components of the reserve vector $\epsilon_{R}$ from (\ref{epsilon_reserved_solution}).
		\item For a sufficiently large iteration number $k$, the values $\mu_{\nu}$ from (\ref{current_lower_bounds}) satisfy the inequalities
			\begin{equation}\label{theorem_inequalities}
				r_{\nu}\mu_{\nu} > 2^{3-\frac{1}{N}}L_{\nu}\sqrt{N+3}, \; 1 \leq \nu \leq m+1. 
			\end{equation}
	\end{enumerate}
	Then, any limit point $\bar{y}$ of the sequence of the trials $\{y^k\}=\big\{y\big(x^k\big)\big\}$ generated by the index algorithm for problem (\ref{problem}) belongs to the set $Y_{\epsilon}$ form (\ref{Y_epsilon}) and satisfies the conditions	
	\begin{eqnarray*}
		\varphi(\bar y) =	\inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \\
		\leq \varphi(y_{\epsilon}).
	\end{eqnarray*}
\end{theorem}
\textbf{Note 1}. The computational scheme of the algorithm is described in the form convenient for  formulating its theoretical properties. The software implementation of this scheme may be made more frugal. For example, according to Step 7, the point $x^{k+1}$ of the next trial will be an internal point of the interval $(x_{t-1}, x_t)$, i.e. $x^{k+1} \in (x_{t-1}, x_t)$. Therefore, the set of trial points $x^1,\ldots, x^k$ (and the related trial results (\ref{trial_result})) may not be ordered  at every step of the algorithm but may be maintained in an ordered state using appropriate data structures.
\\
\textbf{Note 2}. The use of the set of parameters $\epsilon_1,\ldots,\epsilon_m$ (the values of which correspond to the coordinates of the reserve vector $\epsilon_R$) in the index algorithm will significantly reduce the density of the trial sequence in the vicinity of the feasible domain boundaries. However, this raises a question about the particular values of the reserve vector components $\epsilon_R$. In \cite{Strongin2000}, it was proposed to choose the components of the vector $\epsilon_R$ according to the formula 
\begin{equation}\label{epsilon_nu}
	\epsilon_{\nu} = \mu_{\nu}\delta, \; 1 \leq \nu \leq m, 
\end{equation}
where $\mu_{\nu}$ are the current estimates from (\ref{current_lower_bounds}) and $\delta>0$ is a parameter. This approach to choosing the reserve vector components allows one to take into account the properties of different problem functions characterized by their Lipschitz constants. This method for choosing the reserve vector was used in the numerical experiments, the results of which are presented below.
\\
\textbf{Note 3}. The index algorithm for solving the constrained problems is based on the information-statistical algorithm of global search (AGS) for solving unconstrained problems. AGS has been compared repeatedly with both deterministic and nature-inspired global optimization algorithms. The results of the experimental comparison (see, for example, \cite{Sovrasov2019}) show that AGS is superior to many known methods of similar purpose, including DIRECT and DIRECT\textit{l}.

@@

	Let us illustrate the work of the algorithms IA and IA-DL when minimizing the function
\begin{align*}
	\varphi(y_1, y_2) & = -1.5y_1^2\exp{\left\{1-y_1^2-20.25(y_1-y_2)^2\right\}}-\\
	& -\left(0.5(y_1-1)(y_2-1)\right)^4 \\
	& \exp{\left\{2-\left(0.5(y_1-1)\right)^4-(y_2-1)^4\right\}}
\end{align*}
in the range $0 \leq y_1 \leq 4$, $1 \leq y_2 \leq 3$ subject to the constraints
\begin{eqnarray*}
	g_1(y_1, y_2) & = & 0.01 \left( (y_1-2.2)^2+(y_2-1.2)^2-2.25 \right) \leq 0, \\
	g_2(y_1, y_2) & = & 100 \left(1-(y_1-2)^2/1.44+(0.5y_2)^2 \right) \leq 0, \\
	g_3(y_1, y_2) & = & 10 \left( y_2 - 1.5 - 1.5 \sin{\left( 6.283(y_1-1.75) \right)}\right) \leq 0.
\end{eqnarray*}
The feasible domain of this problem is non-connected and consists of three subdomains. The global minimum $\varphi(y^\ast)=-1.489$ is achieved at the point \\ $y^\ast =(0.942, 0.944)$ (the values are given with the accuracy of $10^{-3}$).

\begin{figure}[h]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{stronginC3_IA} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{stronginC3_IA-DL} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Solving a problem by IA (a) and IA-DL (b) methods}
	\label{fig:example_strongin}
\end{figure}

@@

	Let us show that if the conditions of Theorem \ref{theorem:1} are satisfied, the inequality $R_{glob}(j)>0$ will hold. According to the second condition of Theorem \ref{theorem:1} and to the condition (\ref{K_leq_L}), 
\begin{align}
	z_j & = g_{\nu}\left( y(x_j) \right) \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{1/N}, \nonumber \\
	& \nu=\nu(x_j),
\end{align}
\begin{align}
	z_{j-1} &= g_{\nu}\left( y(x_{j-1}) \right) \nonumber \\
	& \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(\bar x - x_{j-1})^{1/N}, \nonumber \\ 
	& \nu=\nu(x_{j-1}).
\end{align}

	Since the point $\bar y = y(\bar x)$ belongs to the set $Y_{\epsilon}$, then
$$
	g_{\nu}\left( y(\bar x) \right) \leq -\epsilon_{\nu}, \; 1\leq\nu\leq m,
$$
$$
	g_{m+1}\left( y(\bar x) \right) \leq z_{m+1}^\ast.
$$

	Then, the characteristic $R_{glob}(j)$ of the interval $[x_{j-1}, x_j], j=j(k)$ will be strictly positive, i.e. $R_{glob}(j)>0$. Indeed, at $\nu=\nu(x_j)=\nu(x_{j-1})$ and the value $r_{\nu}=r_{\nu}^{glob}$
\begin{align*}
	& R(j) = \Delta_i + \frac{(z_j-z_{j-1})^2}{r_{\nu}^2\mu_{\nu}^2\Delta_i} - 2\frac{z_j+z_{j-1}-2z_{\nu}^{\ast}}{r_{\nu}\mu_{\nu}} \\
	& \geq \Delta_i + 4\frac{z_{\nu}^*-\left( g_{\nu}\left( y(\bar x) \right)+L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)\right)}{r_{\nu}\mu_{\nu}} \\
	& = \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^\ast-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
	& \geq \Delta_i\left(1-4\frac{L_{\nu}\sqrt{N+3}\max_{0\leq\alpha\leq1} {\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
	& = \Delta_i\left(1-4\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
\end{align*}
At $\nu=\nu(x_j ) > \nu(x_{j-1})$ and the value $r_{\nu}=r_{\nu}^{glob}$
\begin{align*}
	& R(j) = 2\Delta_i - 4\frac{z_j-z_{\nu}^{\ast}}{r_{\nu}\mu_{\nu}} \\
	& \geq 2\Delta_i+4\frac{z_{\nu}^{\ast}-\left( g_{\nu}\left(y(\bar x)\right)+2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}} \\
	& = 2\Delta_i - 8\frac{L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}}}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
	& \geq 2\Delta_i\left( 1-4\frac{L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right) + 4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
\end{align*}

@@

\begin{figure}[h!]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Simple_59_IA} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Simple_59_IA-DL} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Solving a problem of the “Simple” class by IA (a) and IA-DL (b) methods}
	\label{fig:examples_simple}
\end{figure}

\begin{figure}[h!]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Hard_7_IA} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Hard_7_IA-DL} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Solving a problem of the “Hard” class by IA (a) and IA-DL (b) methods}
	\label{fig:examples_hard}
\end{figure}

@@

\begin{figure}[h!]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklssC_3} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Operating_characteristic_gklshC_3} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Operational characteristics for “Simple” (a) and “Hard” (b) problem classes, $N = 3$}
	\label{fig:operational_characteristics_N=3}
\end{figure}

>>>>

\RequirePackage{fix-cm}

\documentclass[twocolumn]{svjour3}          % twocolumn

\smartqed  % flush right qed marks, e.g. at end of proof
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{lineno,hyperref,amsmath}
\usepackage{multirow}
\usepackage[round]{natbib} 
\usepackage[misc]{ifsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

@@

\begin{abstract}
This paper considers the constrained global optimization problems, in which the functions are of the “black box” type and satisfy the Lipschitz condition. The algorithms for solving the problems of this class require the use of adequate estimates of the a priori unknown Lipschitz constants for the problem functions. A novel approach presented in this paper is based on a simultaneous use of two estimates of the Lipschitz constant: an overestimated and an underestimated one. The upper estimate provides the global convergence whereas the lower one reduces the number of trials necessary to find the global optimizer with the required accuracy. The considered algorithm for solving the constrained problems doesn’t use the ideas of the penalty function method; each constraint of the problem is accounted for separately. The convergence conditions of the proposed algorithm are formulated in the corresponding theorem. The results of the numerical experiments on a series of multiextremal problems with non-convex constraints demonstrating the efficiency of the proposed scheme of dual Lipschitz constant estimates are presented.
\keywords{Global optimization \and Multiextremal problems \and Non-convex constraints \and Lipschitz constant estimates \and Dimensionality reduction \and Numerical methods}
%\subclass{90C26}
\end{abstract}

@@

	It is supposed that the objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_{j}(y)$, $1 \leq j \leq m$, of the constraints satisfy the Lipschitz condition
\begin{align}\label{lipschitz_condition}
	\left|g_j(y')-g_j (y'')\right| \leq L_j \|y'-y''\|, \nonumber \\ 
	y', y''\in D, \; 1\leq j \leq m+1,
\end{align}
with a priori unknown constants $L_j, 1\leq j \leq m+1$, and may be multiextremal.

@@

\begin{figure}
	\center{\includegraphics[width=0.6\linewidth]{Fig1}}
	\caption{$\varepsilon$-reserved solution.}
	\label{fig:eps_reserved_solution}
\end{figure}

@@

	We will consider the initial problem (\ref{problem}) assuming that the problem functions $g_{j}(y)$, $1 \leq j \leq m+1$, may be defined only partly. It means that they are defined and computable only in the sets
\small
\begin{align}\label{D_sets}
	D_1 = D, \; D_{j+1} = \left\{ y \in D_j: \; g_j(y) \leq 0 \right\}, \; 1 \leq j \leq m.
\end{align}
\normalsize
This situation is typical for the optimal design problems when some conditions for the functioning of the modeled object are not met and its other characteristics appear to be undefined. In this approach, the first violation of any constraint indicating that the point $y$ being considered is no longer feasible terminates further computations at the point $y$.

	Many well known approaches to solving the multidimensional global optimization problems are based on either explicit or implicit reduction of multidimensional problems to one-dimensional ones (see, for example, the description of methods using the diagonal partitions \cite{Sergeyev2006} or simplicial partitions \cite{Zilinskas2008}). In this paper, we will use the approach based on the dimensionality reduction with the use of the space-filling curve $y(x)$ (the Peano curve). Continuous unambiguous mapping $y(x)$ projecting the interval [0,1] onto an $N$-dimensional hypercube $D$ allows reducing a multidimensional constrained minimization problem in the domain $D$ to a one-dimensional constrained optimization problem in the interval [0,1]
\small
\begin{align}\label{one_dimensional_problem}
	& \varphi(y(x^\ast)) =  \nonumber \\
	& \min \left\{\varphi(y(x)): x \in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\}.
\end{align}
\normalsize
The issues of the numerical construction of the Peano space filling curve-type mappings and the corresponding theory were described in detail in \cite{Strongin2000,Sergeyev2013}. Note that the theoretical Peano curve $y(x)$ is a limit object and in practice is replaced by its approximation (called \textit{evolvent}) with predefined precision $2^{-m}$, where the number $m>1$ is a parameter (the \textit{density} of the evolvent). The evolvent density is determined by the accuracy of the possible practical implementation of the problem solution found. For  illustration, Fig. \ref{fig:evolvent} (a), (b) shows the projections of the interval [0,1] for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
\begin{figure}
	\begin{minipage}[h]{0.45\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{Fig2} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.45\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{Fig3} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Evolvent for $N=2$ and $m=4$ (a), $m=5$ (b)}
	\label{fig:evolvent}
\end{figure}

@@

	The convergence conditions for the index global search algorithm can be formulated in the form of the following theorem from \cite{Strongin2000}. 
\\
\begin{theorem}\label{theorem:1} Assume that the following conditions are satisfied:
	\begin{enumerate} 
		\item Problem (\ref{problem}) has an $\varepsilon$-reserved solution $y_{\varepsilon}$ from (\ref{epsilon_reserved_solution}).
		\item Functions $g_j(y), 1 \leq j \leq m+1,$ admit Lipschitzian (with the constants $L_j$) extensions $G_{j}(y)$ over the whole feasible domain $D$ from (\ref{D}), i.e.
	$$
		g_j \left( y(x) \right) = G_j \left( y(x) \right), \; x \in Q_j, \; 1 \leq j \leq m+1,
	$$
	where the sets $Q_j$ are from (\ref{Q_intervals}), and $y(x)$ is the curve from (\ref{one_dimensional_problem}).
		\item The parameters of the method $\epsilon_{\nu}, 1 \leq \nu \leq m,$ used in the rule (\ref{z_estimates}) are the corresponding components of the reserve vector $\epsilon_{R}$ from (\ref{epsilon_reserved_solution}).
		\item For a sufficiently large iteration number $k$, the values $\mu_{\nu}$ from (\ref{current_lower_bounds}) satisfy the inequalities
			\begin{equation}\label{theorem_inequalities}
				r_{\nu}\mu_{\nu} > 2^{3-\frac{1}{N}}L_{\nu}\sqrt{N+3}, \; 1 \leq \nu \leq m+1. 
			\end{equation}
	\end{enumerate}
	Then, any limit point $\bar{y}$ of the sequence of the trials $\{y^k\}=\big\{y\big(x^k\big)\big\}$ generated by the index algorithm for problem (\ref{problem}) belongs to the set $Y_{\epsilon}$ form (\ref{Y_epsilon}) and satisfies the conditions
	\begin{eqnarray*}
		\varphi(\bar y) & = &	\inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \\
		& \leq & \varphi(y_{\epsilon}).
	\end{eqnarray*}
\end{theorem}
\textbf{Note 1}. The computational scheme of the algorithm is described in the form convenient for  formulating its theoretical properties. The software implementation of this scheme may be made more frugal. For example, according to Step 7, the point $x^{k+1}$ of the next trial will be an internal point of the interval $(x_{t-1}, x_t)$, i.e. $x^{k+1} \in (x_{t-1}, x_t)$. Therefore, the set of trial points $x^1,\ldots, x^k$ (and the related trial results (\ref{trial_result})) may not be ordered  at every step of the algorithm but may be maintained in an ordered state using appropriate data structures.
\\
\textbf{Note 2}. The use of the set of parameters $\epsilon_1,\ldots,\epsilon_m$ (the values of which correspond to the coordinates of the reserve vector $\epsilon_R$) in the index algorithm will significantly reduce the density of the trial sequence in the vicinity of the feasible domain boundaries. However, this raises a question about the particular values of the reserve vector components $\epsilon_R$. In \cite{Strongin2000}, it was proposed to choose the components of the vector $\epsilon_R$ according to the formula 
\begin{equation}\label{epsilon_nu}
	\epsilon_{\nu} = \mu_{\nu}\delta, \; 1 \leq \nu \leq m, 
\end{equation}
where $\mu_{\nu}$ are the current estimates from (\ref{current_lower_bounds}) and $\delta>0$ is a parameter. This approach to choosing the reserve vector components allows one to take into account the properties of different problem functions characterized by their Lipschitz constants. This method for choosing the reserve vector was used in the numerical experiments, the results of which are presented below.
\\
\textbf{Note 3}. The index algorithm for solving the constrained problems is based on the information-statistical algorithm of global search (AGS) for solving unconstrained problems. AGS has been compared repeatedly with both deterministic and nature-inspired global optimization algorithms. The results of the experimental comparison (see, for example, \cite{Sovrasov2019}) show that AGS is superior to many known methods of similar purpose, including DIRECT and DIRECT\textit{l}.

@@

	Let us illustrate the work of the algorithms IA and IA-DL when minimizing the function
\small
\begin{align*}
	&\varphi(y_1, y_2) = -1.5y_1^2\exp{\left\{1-y_1^2-20.25(y_1-y_2)^2\right\}}-\\
	& -\left(0.5(y_1-1)(y_2-1)\right)^4 \exp{\left\{2-\left(0.5(y_1-1)\right)^4-(y_2-1)^4\right\}}
\end{align*}
\normalsize
in the range $0 \leq y_1 \leq 4$, $1 \leq y_2 \leq 3$ subject to the constraints
\small
\begin{eqnarray*}
	g_1(y_1, y_2) & = & 0.01 \left( (y_1-2.2)^2+(y_2-1.2)^2-2.25 \right) \leq 0, \\
	g_2(y_1, y_2) & = & 100 \left(1-(y_1-2)^2/1.44+(0.5y_2)^2 \right) \leq 0, \\
	g_3(y_1, y_2) & = & 10 \left( y_2 - 1.5 - 1.5 \sin{\left( 6.283(y_1-1.75) \right)}\right) \leq 0.
\end{eqnarray*}
\normalsize
The feasible domain of this problem is non-connected and consists of three subdomains. The global minimum $\varphi(y^\ast)=-1.489$ is achieved at the point \\ $y^\ast =(0.942, 0.944)$ (the values are given with the accuracy of $10^{-3}$).

\begin{figure}[h]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Fig4} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Fig5} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Solving a problem by IA (a) and IA-DL (b) methods}
	\label{fig:example_strongin}
\end{figure}

@@

	Let us show that if the conditions of Theorem \ref{theorem:1} are satisfied, the inequality $R_{glob}(j)>0$ will hold. According to the second condition of Theorem \ref{theorem:1} and to the condition (\ref{K_leq_L}), 
\begin{align}
	& z_j = g_{\nu}\left( y(x_j) \right) \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{1/N}, \nonumber \\
	& \nu = \nu(x_j),
\end{align}
\begin{align}
	z_{j-1} & = g_{\nu}\left( y(x_{j-1}) \right) \nonumber \\
	& \leq g_{\nu}\left( y(\bar x) \right) + 2L_{\nu}\sqrt{N+3}(\bar x - x_{j-1})^{1/N}, \nonumber \\
	& \nu = \nu(x_{j-1}).
\end{align}

	Since the point $\bar y = y(\bar x)$ belongs to the set $Y_{\epsilon}$, then
\begin{equation*}
	g_{\nu}\left( y(\bar x) \right) \leq -\epsilon_{\nu}, \; 1\leq\nu\leq m,
\end{equation*}
\begin{equation*}
	g_{m+1}\left( y(\bar x) \right) \leq z_{m+1}^\ast.
\end{equation*}

	Then, the characteristic $R_{glob}(j)$ of the interval \\
$[x_{j-1}, x_j]$, $j=j(k)$ will be strictly positive, i.e. \\
$R_{glob}(j)>0$. Indeed, at $\nu=\nu(x_j)=\nu(x_{j-1})$ and the value $r_{\nu}=r_{\nu}^{glob}$
\small
\begin{align*}
	& R(j) = \Delta_i + \frac{(z_j-z_{j-1})^2}{r_{\nu}^2\mu_{\nu}^2\Delta_i} - 2\frac{z_j+z_{j-1}-2z_{\nu}^{\ast}}{r_{\nu}\mu_{\nu}} \geq \Delta_i + \\ 
	& + 4\frac{z_{\nu}^*-\left( g_{\nu}\left( y(\bar x) \right)+L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)\right)}{r_{\nu}\mu_{\nu}} \\
	& = \Delta_i-4\frac{L_{\nu}\sqrt{N+3}\left( (x_j-\bar x)^{\frac{1}{N}}+(\bar x - x_{j-1})^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}} + \\
	& + 4\frac{z_{\nu}^\ast-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
	& \geq \Delta_i\left(1-4\frac{L_{\nu}\sqrt{N+3}\max_{0\leq\alpha\leq1} {\left( \alpha^{\frac{1}{N}}+(1-\alpha)^{\frac{1}{N}} \right)}}{r_{\nu}\mu_{\nu}} \right) + \\
	& + 4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
	& = \Delta_i\left(1-4\frac{2^{3-1/N}L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right)+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
\end{align*}
\normalsize
At $\nu=\nu(x_j ) > \nu(x_{j-1})$ and the value $r_{\nu}=r_{\nu}^{glob}$
\begin{align*}
	& R(j) = 2\Delta_i - 4\frac{z_j-z_{\nu}^{\ast}}{r_{\nu}\mu_{\nu}} \\
	& \geq 2\Delta_i+4\frac{z_{\nu}^{\ast}-\left( g_{\nu}\left(y(\bar x)\right)+2L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}} \right)}{r_{\nu}\mu_{\nu}} \\
	& = 2\Delta_i - 8\frac{L_{\nu}\sqrt{N+3}(x_j-\bar x)^{\frac{1}{N}}}{r_{\nu}\mu_{\nu}}+4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} \\
	& \geq 2\Delta_i\left( 1-4\frac{L_{\nu}\sqrt{N+3}}{r_{\nu}\mu_{\nu}} \right) + 4\frac{z_{\nu}^{\ast}-g_{\nu}\left( y(\bar x) \right)}{r_{\nu}\mu_{\nu}} > 0.
\end{align*}

@@

\begin{figure}[h!]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Fig6} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Fig7} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Solving a problem of the “Simple” class by IA (a) and IA-DL (b) methods}
	\label{fig:examples_simple}
\end{figure}

\begin{figure}[h!]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Fig8} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Fig9} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Solving a problem of the “Hard” class by IA (a) and IA-DL (b) methods}
	\label{fig:examples_hard}
\end{figure}

@@

\begin{figure}[h!]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Fig10} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Fig11} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Operational characteristics for “Simple” (a) and “Hard” (b) problem classes, $N = 3$}
	\label{fig:operational_characteristics_N=3}
\end{figure}

\begin{figure}[h!]
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Fig12} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.48\linewidth}
		\center{\includegraphics[width=\linewidth]{Fig13} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Operational characteristics for “Simple” (a) and “Hard” (b) problem classes, $N = 5$}
	\label{fig:operational_characteristics_N=5}
\end{figure}
